---
title: A Hitchhiker's Guide to Reproducible Research in R
author:
  - name: Aaron Peikert
    affil: 1, *
    orcid: 0000-0001-7813-818X
  - name: Caspar J. Van Lissa
    affil: 2, 3
    orcid: 0000-0002-0808-5024
  - name: Andreas M. Brandmaier
    affil: 1, 4
    orcid: 0000-0001-8765-6982
affiliation:
  - num: 1
    address: |
      Center for Lifespan Psychology---Max Planck Institute for Human Development
      Lentzeallee 94, 14195 Berlin, Germany
  - num: 2
    address: |
      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
  - num: 3
    address: |
      Open Science Community Utrecht, Utrecht, Netherlands
  - num: 4
    address: |
      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
      Berlin, Germany and London, UK
# firstnote to eighthnote
correspondence: |
  peikert@mpib-berlin.mpg.de
journal: psych
type: tutorial
status: submit
bibliography: temp.bib
simplesummary: |
  A Simple summary goes here.
abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  The authors received no financial support for the research, authorship, and/or publication of this article.
authorcontributions: |
  Aaron Peikert took the lead in writing and provided the initial draft of the manuscript.
  Andreas Brandmaier and Caspar J. Van Lissa contributed further ideas, critical feedback, and revisions of the original manuscript.
  Furthermore, we would like to thank Maximilian Stefan Ernst (not an author) for his contributions to the code for the simulation study.
conflictsofinterest: |
  The authors declare no conflict of interest.
abbreviations:
  - short: PAC
    long: Preregistration as 
  - short: Gb
    long: Gigabyte
  - short: Kb
    long: Kilobyte
  - short: CRAN
    long: Comprehensive R Archive Network
repro:
  packages:
    - tidyverse
    - usethis
    - gert
    - aaronpeikert/repro@c3cae1f
    - here
    - rstudio/webshot2@f62e743
    - targets
    - renv
    - slider
    - patchwork
    - knitr
    - pander
    - lavaan
    - furrr
    - future.batchtools
    - rticles
    - moments
    - report
  scripts:
    - R/simulation.R
    - R/simulation_funs.R
    - R/link.R
  data:
    - data/simulation_results.csv
output:
  bookdown::markdown_document2:
    base_format: rticles::mdpi_article
header-includes:
   - \usepackage[labelformat=empty]{caption}
   - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, numbers=left}
---

```{r setup, include=FALSE}
library(repro)
automate_load_packages()
automate_load_scripts()
source(here::here("R", "link.R"))
source(here::here("R", "simulation_funs.R"))
```

```{r echo=FALSE}
# define a print method for objects of the class data.frame
knit_print.data.frame = function(x, ...) {
  res = paste(c('', '', kable(x)), collapse = '\n')
  asis_output(res)
}
# register the method
registerS3method("knit_print", "data.frame", knit_print.data.frame)
```

# Introduction

Scientists increasingly strive to make research data, materials, and analysis code openly available.
Sharing these types of digital research artifacts has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
To some extent these potential benefits are contingent on whether these digital research artifacts are reproducible. 
Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
The credibility of empirical research hinges on its objectivity. Objectivity in this context means "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper]. Therefore, reproducibility has long been considered integral to empirical research.
<!---
You tend to use quite long sentences with many subclauses. While this has a long tradition in the german language, it is considered bad style in english. Also it often leads to minor mistakes like dangling modifiers.
-->
Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
We now have the technology to make all artefacts related to a research project available so that a research project can be reproduced with minimal effort.

This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
We pursue two aims.
First, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
This workflow builds on several technical solutions originating in the software engineering community. 
We discuss how these solution address four common threats to reproducibility [see @Peikert2019 for details] and form a coherent workflow.
We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
Second, we discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including the Introduction, Methods, and Results sections.
The resulting dynamic document closely resembles the final manuscript, but uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
Once the empirical data are analyzed, the Results are updated, and the Discussion is written.

[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalized to inter-subjective criticism (Footnote *1, p. 22). Somewhat confusingly, @popperLogicScientificDiscovery2002 uses the term "Reproducibility" for what we would call "Replication".

Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
<!---
I am not shure about the word artifact in general. Is it common terminology?
-->
However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
Out of 35 published articles with open research artifacts, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
If research artifacts are available but not reproducible, their added value is limited.

This tutorial therefore demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
An easy-to-use standard, on the other hand, is more likely to be adopted.
Our approach is to promote broad adoption of easy-to-use practices (a "low threshold"), and ensure that these practices are compatible with more complex rigorous solutions (a "high ceiling").
As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
The first part of this tutorial outlines a reproducibility standard, and introduces the underlying theoretical concepts and software solutions.
The second part is a tutorial that demonstrates how this standard can be easily adopted with the help of the R package `repro`.

We have structured the tutorial with a *learning by doing* approach in mind, such that readers can follow along on their own computers.
We explicitly encourage readers to try out all R commands for themselves.
Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).

# Concepts and Software Solutions<!--CJ Some inconsistencies:
1) From the end of the previous paragraph, I would expect a description of the "rigorous reproducibility standard" here.
2) The paragraph is labeled "Concepts and Software SOlutions, but it begins with a list of threats to reproducibility. Threats are not concepts nor software solutions.
3) After the list of threats, we get a description of the "rigorous reproducibility standard". That is what I would expect based on the preceding paragraph. Note that the "rigorous reproducibility standard" is also not a concept, nor a software solution. Consider re-naming this section.-->

From our own experience with various research projects, we identified the following common threats to reproducibility: <!-- CJ This still requires attention! How did you identify these causes? Is this exhaustive?-->

1.	**Multiple inconsistent versions of code, data, or both**; for example
  - the data set could have changed over time because outliers were removed at a later stage or an item was later recoded
  - the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
It may then be unclear which version of code and data was used to produce some reported set of results.
2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
<!-- CJ 5. Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
<!--AB Caspar's 5th point is important. In my mind, this best fits to point #3, which adressed an insufficient level of coding/automation/documentation of the necessary steps taken. Can we merge this into one new point #3? I will commit a suggested change of point #3 here (see above);.-->

We have developed a workflow that achieves long-term and cross-platform computational reproducibility of scientific data analyses. It leverages established tools and practices from software engineering and rests on four pillars that address the aforementioned causes of non-reproducibility: [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph. The opening paragraph should at least:
1) Introduce the topic area
2) Introduce the problem/knowledge gap
3) Introduce how you will fill this knowledge gap-->

1. Version control
2. Dynamic document generation
3. Dependency tracking
4. Software management

The remainder of this section briefly explains why each of these four building blocks is needed and their role in ensuring reproducibility.

**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the development phase of a research project.
Version control allows a clear link between which results were generated by which version of code and data.
We recommend using **Git** for version control, because of its widespread adoption in the R community.

Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
Git is built around snapshots that represent the project state at a given point in time.
Those snapshots are called "commits" and work like a "save" action.
Changes to files are collected in a _commit_.
Ideally, each commit has a message that succinctly describes these changes.
It is good practice to make commits for concrete milestones, e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3".
This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
Each commit "knows" its ancestor, and they form a timeline together.
The entirety of commits (i.e. the version-controlled project) is called a repository.
In Git, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
Git has additional features beyond basic version control, such as branches to facilitate collaboration.
@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git for research.
@bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
Collaborating via Git requires the repository to be uploaded somewhere.
We recommend GitHub to host Git repositories because of its popularity among R users.
GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists everyday work, but they are not central to achieving reproducibility.

Second, we rely on **dynamic document generation**.
The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
Typically, results are manually copied and pasted from one software another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].

Dynamic document generation integrates both steps.
Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allows interested readers deeper understanding [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
**R Markdown** uses Markdown for writing the text and R (or other programming languages) for the analysis.
Markdown is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
This way, Markdown does not need any specialized software for editing.
It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
This allows the creation of formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.

One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.

To automatically resolve dependencies between project files, we rely on **dependency tracking**.
In essence, researchers provide a collection of computational instructions (_recipes_) that describe how ingredients are processed to create intermediate products and, in the end, a final product.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).

We recommend using **Make** for dependency tracking because it is language independent.
The following hypothetical example illustrates the utility of Make.
Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
A Makefile for this project could look like this:

```{bash, eval=FALSE}
manuscript.pdf: manuscript.Rmd simulated_data.csv
  Rscript -e 'rmarkdown::render("manuscript.Rmd")'

simulated_data.csv: simulate.R
  Rscript -e 'source("simulate.R")'
```

There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
Each target is followed by a colon and a list of requirements.
If a requirement is newer than the target, the recipe will be executed to rebuild the target.
If a requirement does not exist, Make uses a recipe to build the requirement first before building the target.
Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
This ensures that the simulation is always run before the manuscript is built, and that the manuscript is rebuilt if the simulation code was changed.
Make therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.

```{r, include=FALSE}
container_size <- as.numeric(fs::file_size("reprotutorial.sif"))/1024^3
```

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

**Docker** is an example of containerization.
It provides all software used in an analysis, including the operating system itself.
(The operating system level is important because some functionality may be passed to software layers beneath the programming language, for example random number generators and linear algebra libraries.)
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Researchers therefore have the opportunity to use the exact same software setup that was used by the original authors.
Such a snapshot of the software stack is called an "image".
Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
A second strategy is to save a container recipe, a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend storing a complete image of the software environment besides the Dockerfile.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)); dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).    
While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
To remove this hurdle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.

```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown). Git tracks changes to the project over time. Make manages dependencies among the files. Docker provides a container in which the final report is built using dynamic document generation in R Markdown. Reproduced from @Peikert2019.  "}
# file gets downloaded in Makefile
knitr::include_graphics(here::here("images/nutshell.svg"), auto_pdf = TRUE)
```

# Creating reproducible research projects

The above mentioned tools require expertise that is beyond the traditional curriculum of researchers.
To aid researchers adopting these tools we rely on and have developed helpers that wrap more complicated (and powerful) tools.
These helpers wrap standard tools but provide feedback about what the computer does and what the user should do in layman's terms and are accessible from within R.
These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.

Before we can start creating reproducible research projects, we need a computer with a working R Studio installation. 
Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so it can assist you following the tutorial:

```{r, eval=FALSE}
# repro is not on CRAN yet
install.packages('repro', repos = 'https://aaronpeikert.r-universe.dev')
```

An example of the assistance that `repro` provides are the check functions.
They verify that you indeed have installed and setup the required pieces of software for this workflow (we use the double-colon operator to access functions from the `repro` namespace to make clear which functions are native `repro` functions):

```{r}
# `package::function()` → use function from package without `library(package)`
repro::check_git()
repro::check_make()
repro::check_docker()
```

They detect the users' system state and explain if further action is needed.
For example, the following happens if you were a Windows user who does not have Git installed:

```{r, include=FALSE}
opts <- options()
options(repro.os = "windows",
        repro.git = FALSE)
```

```{r}
check_git()
```

```{r, include=FALSE}
options(opts)
```

The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
This tutorial describes steps that we expect to apply to all users; the messages from the packages are adjusted to your specific situation (that is, operating system and the specific problem of missing software) and complement the tutorial in this way.
Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
If necessary, follow the instructions of repro to install them until the check functions are all fine.

The first concept we want to introduce is version control with Git.
Git tracks files within one project folder.
We start by creating such a project folder with RStudio by clicking the menu item:

> File → New Project... → New Directory → Example Repro Template

This creates a project containing an example analysis, but you may use any other template or existing R Project to turn into a reproducible research project using the functionality of `repro`.

You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore` file.
You can edit the `.gitignore` file directly or with this command:

```{r, eval=FALSE}
usethis::use_git_ignore("private.md")
```

Now the file `private.md` will not be committed to Git and will not be made public.

After you have excluded sensitive files (like not yet anonymized data files), you can activate Git with:

```{r, eval=FALSE}
usethis::use_git()
```

For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
The easiest way to do this is to use the RStudio Git pane.
Click on the empty box next to the file you want to stage, and when a tick appears, which means that the file is staged.
After you have staged all files you would like to, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
<!---
Maybe first click on the commit button and then stage, because in the bigger window you can browse the changes.
-->

The files you created and the changes you made still have not left your computer.
All snapshots are stored in a local repository in your project folder.
To back up and share the files, you can publicly upload those changes to a public repository.
To synchronize your local repository with a remote repository from the GitHub service, issue the following command:

```{r, eval=FALSE}
usethis::use_github()
```

Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the R console.

If you are not yet ready to share your project publicly but you want to synchronize your local repository to a remote repository (for example, as a simple backup service), use:

```{r, eval=FALSE}
usethis::use_github(private = TRUE)
```

Now that you have created a project that is under version control, we continue with adding dynamic document generation.
A dynamic document has three elements:

1. Text
2. Executable code
3. Metadata

The text of an R Markdown is formatted by Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
The following example serves to illustrate the Markdown syntax.
It shows how to create a heading, a word in bold font, and a list of several items in Markdown:

```markdown
<!--this is a Markdown file -->
# Heading (level 1)

Normal text.
Important **word** in bold.

## Subheading (level 2)

To do list:

* do research
* do more research
* spend time with family and friends
```

This type of markup for formatting, instead of the more visual approach of e.g. Microsoft Word, is compatible with many output formats and is easier to put under version control.
Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
The visual editor has the additional benefit of promoting best practices (for example, each sentence should be written on a new line, which eases tracking changes across versions).
Among other features, it eases the generation of citations and cross references.

Now that you are familiar with the basics of writing text in Markdown, we turn our attention to including code into the text.
Code is separated by three backticks (and the programming language in curly braces):

````markdown
This is normal text, written in Markdown.

`r ''````{r}
# this is R code
1 + 1
```
````

The results of your code are dynamically inserted into the document.
Dynamically means that whenever you turn the R Markdown into its output format the code is executed again and if you changed something the results are updated.
<!---
you did not explain how to knit and to choose an output format
-->
The output document is then static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
This static document is somewhat decoupled from the R Markdown and the rest of the code that did create it.
It is quite possible that several slightly different versions are therefore circulating.
We therefore recommend to connect the static document to the version control system by dynamically including the identifier of the Git commit, called hash, in the document.
The `repro` package comes with the function `repro::current_hash()` for that purpose.
This document was created from the commit with the hash ``r repro::current_hash()``.

Now that you know how to write text and R code in an R Markdown, you need to know about metadata, like the title or the output format. Metadata are placed at the beginning of the document and are separated by three dashes from the document body. The following example is a full markdown document where the the metadata (the "header") is in lines 1--6.

````markdown
---
title: A Hitchhiker's Guide to Reproducible Research in R
author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
abstract: A tutorial about doing the same thing more than once.
output: html_document
---

# Introduction

Important for reproducibility:

1. Version Control
2. Dynamic Document Creation
3. Dependency Tracking
4. Software Management

`r ''````{r}
# this is R code
t.test(extra ~ group, data = sleep)
```
````

Most metadata fields are self-explanatory (like the author field). 
Sometimes, R packages may extend the metadata fields for their own purposes.
Certain output formats may require specific  metadata. 
For example, the `rticles`-package, which offers templates for authoring scientific manuscripts, comes with extra fields that are described in the package documentation and offer extra options for flexible formatting of documents.
R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
The `repro` package uses the metadata for hints to make the research project reproducible.
Researchers should list all dependent resources of their research project in the metadata, that is, R scripts, data files, and external packages, in the following format (see everything below the line *repro:*):

```markdown
---
title: A Hitchhiker's Guide to Reproducible Research in R
author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
abstract: A paper about recalculating the same thing more than once.
output: html_document
repro:
  scripts:
    - R/load.R
  data:
    - data/mtcars.csv
  packages:
    - tidyverse
    - usethis
    - gert
---
```

This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
These two files together form the basis for consistency within a research project and consistency across different systems.
<!---
What is consistency?
-->
At this point, we would like to emphasize that a prerequisite for reproducibility is that every step taken in an analysis can be **automatically** reproduced.
This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).

In everyday work, the simplest solution is to now call the function `repro::automate()` which creates a Makefile and a Dockerfile from all R Markdown files in the project based on the respective repro metadata entries.
Every time you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git. <!--AB: why is this addition not automated? Feels strange given that this is our recommended best practice?!-->

When you or someone else who received your project would like to reproduce it, they can ask repro to explain how this should be done:

```{r}
repro::reproduce()
```

If you feel uncomfortable using the terminal directly, you can sent the command to the terminal from within R:

```{r, eval=FALSE}
system(reproduce())
```

Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
These projects will remain reproducible irrespective of whether `repro` will be maintained and remain accessible in future.
Users do not have to have repro installed to reproduce a project, in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed. 
The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
The commands can be pasted into a terminal and will reproduce the project in many cases.
<!--AB: I said this before. I think it really should be possible to execute this command from within R. Never underestimate the troubles people can have with the terminal. I think we discussed this before. Please add a function repro::rerun(execute=TRUE) or -- if we rename it -- repro::reproduce(execute=TRUE). -->

# Advanced Features

Some projects, however, are not so easy to make reproducible.
As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
These simplifications are possible because of two restrictions.
Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g. raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
All other users may safely skip this section and return if they need to.

The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
Other software that users might need like other programming languages and not preinstalled LaTeX packages have to be added manually.
We plan to add support for commonly used ways to install software beyond R packages, e.g. for system libraries (via the Ubuntu package manager `apt`), LaTeX packages (via the Tex Live package manager `tlmgr`), and python packages (via the python package manager `pip`).
The second limitation is about dependencies.
Make can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
If B is missing in this example, Make would know how to recreate it from C and D.
These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
Users therefore have to either "flatten" the dependency structure by simply stating that A depends on B, C and D, thereby leaving out important information, or express the dependencies directly within the Makefile.

The following section explains how to overcome these limitations while still relying on the automations provided by `repro`.
Lifting these limitations requires the user to interact more directly with Make or Docker.
To satisfy more complicated requirements users need to understand how repro utilizes Make and Docker internally.
<!-- I am pretty shure you are saying that users should satisfy more complicated requirements ^^ -->

Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`. It consists of two processing steps:
First it recreates the virtual software environment (Docker) and then it executes computational recipes in the virtual software environment (Make).
The first step is done by the command `make docker`, it will trigger Make to build the target called `docker`.
The recipe for this target builds an image from the Dockerfile of the repository.
The `&&` concatenates both commands and only runs the second command if the first was successful.
Therefore, the computational steps are only executed when the software environment is set up.
The second step executes the actual reproduction and is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy parts.
First, a call to `make` without any explicit target will build the Make target `all`.
Second, the flag `-B` means that Make will consider all dependencies as outdated and hence will rebuild everything.
Third, repro constructs Make targets so that if you supply `DOCKER=TRUE` they are executed within the Docker image of the project.

This interplay between Docker and Make is relatively complicated because it resembles a chicken or egg problem.
We have computational steps (Make) which depend on the software environment (Docker) for which we again have computational steps that create it.
Users only require a deeper understanding of this interdependence when they either want to have more complex computational recipes than rendering an R Markdown or require other software than R packages.

Users can have full control over the software installed within the image of the project.
`repro` creates three Dockerfiles inside the `.repro` directory.
The file `.repro/Dockerfile_base` contains information about the base image on which all the remaining software is installed.
By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
These contain (among other software) the packages `tidyverse`, `rmarkdown` and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
Users can choose which R version they want to have inside the container by changing the version number in line 1 to the desired R version number.
By default the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
The build date is used to install packages in the version available on CRAN on this specific date and can also be changed, by default it is set to the date on which `repro::automate()` was called the first time. 
This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
Below, you see the Docker base file we used to create this manuscript:

```{r, results='asis', echo=FALSE}
cat(
  "```bash",
  readLines(here::here(".repro", "Dockerfile_base"))[1:3],
  "```",
  sep = "\n"
)
```

The file `.repro/Dockerfile_manual` is empty by default and can be used to add further dependencies outside of R, like system libraries or external software. 
Remember, the software environment of the Docker image is completely separate from software installed on your computer.
This separation is excellent for reproducibility, but requires you to install software on an operating system that may not be familiar to you.
The images supplied by [@boettigerIntroductionRockerDocker2017] are based on Ubuntu, and hence software has to be installed in a way that is compatible with Ubuntu.
The most convenient way to install software on Ubuntu is through its package manager `apt`.
If the following snipped is added to `.repro/Dockerfile_manual` the Docker image will have Python installed.

```bash
RUN apt-get update && apt-get install -y python3
```

The last Dockerfile, `.repro/Dockerfile_packages`, is automatically generated from the metadata provided and must not be manually changed.
Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.

The separation into several Dockerfiles is a design choice to provide a clear separation between automatically generated parts based on metadata provided and manually altered parts of the Dockerfile.
Docker eventually requires a single Dockerfile to run, so `repro::automate()` simply concatenates the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the R project.

With this approach, users of `repro` can not only build complex software environments but they can also implement complex file dependencies.
The standard `repro` metadata only makes sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
If you can formulate the creation of dependencies in terms of computational steps, e.g. the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the Makefile.
The Makefile that `repro` creates is only a template and you are free to change it.
However, make sure to never remove the following two lines:

```bash
include .repro/Makefile_Rmds
include .repro/Makefile_Docker
```

The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the R Markdown files.
This file should not be altered manually.
If you are not satisfied with the automatically generated target simply provide an alternative target in the main `Makefile`.
Targets in the main `Makefile` take precedent.

The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but usually should not modify.
This Makefile coordinates the interplay between Make and Docker and contains targets for building (with `make docker`) and saving (with `make save-docker`) the Docker image.
Additionally, it provides facilities to execute commands within the container.
If you write a computational recipe for a target it will be by default evaluated using the locally installed software.
To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first Make example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l. 2:
```bash
simulated_data.csv: R/simulate.R
  $(RUN1) Rscript -e 'source("R/simulate.R")' $(RUN1)
```

If users execute this in the terminal:

```bash
make simulated_data.csv
```
It behaves exactly as the in the first Make example, the script `R/simulate.R` is run using the locally installed R.
Because this translates simply to:

```{r, echo=FALSE, results='asis'}
cat("```bash",
  stringi::stri_wrap(system2("make", " --dry-run simulated_data.csv", stdout = TRUE)),
  "```", sep = "\n")
```

But if users use

```bash
make DOCKER=TRUE simulated_data.csv
```

it is evaluated within the Docker container using the software within it and not the locally installed R version:

```{r, echo=FALSE, results='asis'}
cat("```bash",
  stringi::stri_wrap(system2("make", " --dry-run DOCKER=TRUE simulated_data.csv", stdout = TRUE)),
  "```", sep = "\n")
```

To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in form of Make) and software management (using Docker) without the necessity to learn both languages.
Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automations.

# Preregistration as Code 

Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
For a priori planned analyses, preregistration should reduce "researcher degrees of freedom", the number of decisions researchers can make that could influence the results.
However, some problems described in the literature make it questionable whether preregistration can achieve its goals.
Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
The current best practice to ensure that the preregistration is comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
The structure these templates provide can more effectively reduce researcher degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
Nevertheless, these templates can not remove all researcher degrees of freedom because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.

Our proposed solution is simple:
Instead of describing the analysis, you write and publish the code to conduct the analysis before you gather data.
<!-- I am not shure if it is good to directly talk to the reader in this form -->
We term this *preregistration as code (PAC)*.
PAC has the potential to eliminate all researcher degrees of freedom, even those standard preregistration can not eliminate.

The idea of submitting code as part of a preregistration is not new.
A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`), we found close to a dozen published articles that included some form of script as part of their preregistration.

We believe, however, that combining ideas from reproducibility and preregistration can turn this idea into a tool with tangible benefits for the research community.

## Why are PAC and reproducibility related? 

First, preregistrations are plagued by non-reproducibility like any other research outcome.
@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis, but from these, only 34 (20%) provided enough information to reproduce the power analysis.
Even researchers who have gone to great lengths in submitting a script in their preregistration sometimes inexplicably fail to reproduce their results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).

Second, although it has been almost a decade since @steegenMeasuringCrowdAgain2014 have preregistered their code, the idea of publishing code before analyzing data has gained not much traction.
<!-- This is not a reason why PAC and reproducibility are related -->
This could change when combining the here proposed workflow and PAC.
We want to extend PAC from preregistering code to preregistering *the whole manuscript as a dynamic document*.
The idea is as follows:
For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
In this document, the results are initially based on simulated data that mimics the expected output.

Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
We have already explained the need for reproducibility standards in the process of writing such manuscripts and it is therefore only natural to extend these arguments to preregistration.

## Advantages of PAC over traditional preregistration

We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
First, a PAC removes any ambiguity regarding the translation of a verbal analysis plan into code.
Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods or an analysis may perform automated variable selection mechanisms before running the final model.
Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
We believe PAC reveals missing steps or ambiguous decisions. 
It can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.

## Benefits of PAC for study planning

Leaving meta-scientific benefits aside, a preregistration is also a tool for study planning.
To use PAC for study planning, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
Then they should translate their method section into executable computer code and simulate data that has the same structure as the data they expect to obtain from their experiments.

We consider three domains where this procedure can help planning a study.
First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
Third, the plan should be efficient; it should only require the necessary amount of resources.

### Comprehensiveness

Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
<!-- I dont understand this sentence -->
One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.

Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
While relatively uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
When simulating data, researchers have to think about and address data-analytic problems on a conceptual level before making an effort to acquire data [@axelrodAdvancingArtSimulation1997].
It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
Writing incomplete code is technically impossible, or at least researchers would notice if something is omitted (e.g. if they forgot to aggregate the data).


### Effectiveness

With simulated data, researchers can check if the analysis recovers the known parameters of the simulation.
When an analysis fails to recover the correct values in idealized conditions with big samples, researchers should not expect this analysis to be effective under realistic conditions.
Because the simulation captures the researchers' expectations about the real world, it is the best way to assess the effectiveness of the analysis without actually gathering the data.
<!-- I would maybe say something about why the simulated data captures researchers expectations. I think for most people data simulation is a novel topic, and even more so the idea that complicated assumptions about the real world (skewness, factor structure of a test, violated measurement invariance) can be incorporated into the simulation -->

### Efficiency

A single simulated dataset already helps to ensure that the plan is viable under certain circumstances.
However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
Monte Carlo simulation lets researchers navigate several trade-offs regarding their resources.
In the simplest case, they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].

## What you need to know before writing a PAC

To write a PAC, researchers should start by drafting a manuscript in the form of an R Markdown.
When they have completed the method section, they should begin writing code.
We suggest writing three functions for each hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
This function should receive the data and derive relevant results from it.

```{r, eval=FALSE}
planned_analyis <- function(data){
  # 1. preprocess e.g. with `rowMeans(data)`
  # 2. conduct analysis e.g. with `t.test()`
  # 3. return the results
}
```

Because it is difficult to draft such a function without interactively testing it, the researcher should then write a function for simulating data.

```{r, eval=FALSE}
simulate_data <- function(n, effect_size, measurement_error, etc.){
  # 1. warn users that the results are mocked
  # 2. draw `n` samples with `effect_size`
  # 3. Further assumptions can be passed to this function, like measurement error, skewness, etc.
  # 4. format and return in expected data format
}
```

As soon as they have written a first draft of `planned_anaylsis()` and `simulate_data()` they can interactively improve both functions, e.g. till `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
The goal is that the output of `simulate_data()` works as input to the function `planned_analysis()`.

This process can be tedious and complicated.
If it proves infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
Of course, researchers must not analyze the data before they write the preregistration.
However, they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.

When the researchers are satisfied with the function `planned_analysis()`, they can think about how they would like to report the analysis results via tables, plots, and text.
The implementation of this reporting should be in the function `report_analysis()`.
<!-- Inconsistent naming of the function report_analysis() !! -->

```{r, eval=FALSE}
report_results <- function(results){
  # 1. create markdown tables from results
  # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
  # 3. visualize results
  # 4. return results section formatted in markdown
}
```

This function should again accept the output of `planned_analysis()` as input.
The output of this function should be formatted in Markdown.
Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g. it quantifies the size of an effect).

Ideally, these three functions can be composed to create a mock results section, e.g. when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` they output a full results section.

To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data) and write the discussion.

## An example of Preregistration as Code

We have made the argument that PAC has significant advantages over classical preregistration and have outlined how to implement it, but some may be sceptical if it is feasible.
To address this concern and help researchers implementing PAC themselves, we have conducted a PAC that can serve as an example.
We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:

> "Is there a mean difference in the personality trait 'Machiavellism' between self-identified females and males?"

Again, we propose a preregistration format that closely resembles a classical journal article, but utilizes simulated data and dynamic document generation to serve simultaneously as preregistration and final manuscript.
The complete example is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
In this section, we show excerpts of this example formatted in monospace and provide some commentary.

```{r, echo=FALSE}
show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
  stopifnot(length(x) == 1L)
  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
  if(!isFALSE(fence)){
    if(isTRUE(fence))fence <- "markdown"
    fence <- str_c("````", fence)
    x <-
      str_c(fence,
            x,
            "````",
            sep = "\n",
            collapse = "\n")
  }
  if(cat){
    cat(x)
    return(invisible(x))
  } else {
    return(x)
  }
}
show_prereg <-
  function(section,
           file = "preregistration.Rmd",
           wrap = TRUE,
           cat = TRUE,
           fence = TRUE) {
    prereg <- readr::read_lines(here::here(file))
    start <- stringr::str_c("<!--begin-", section, "-->")
    end <- stringr::str_c("<!--end-", section, "-->")
    start_pos <- which(prereg == start)
    end_pos <- which(prereg == end)
    if (length(start_pos) != 1L || length(end_pos) != 1L)
      stop("No section '", section, "' found.")
    excerpt <-
      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
    show(excerpt, wrap = wrap, cat = cat, fence = fence)
  }
```

As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary information and literature to understand the rest of the paper.
This example is drastically shortened for illustration purposes:

```{r, echo=FALSE, results='asis'}
show_prereg("theory")
```

After researchers have provided the research question, they proceed to explain how they want to study it.
For simplicity, we will use already published data that we have not yet analyzed:

```{r, echo=FALSE, results='asis'}
show_prereg("method")
```

We choose the following statistical procedure because many researchers may be familiar with it[^caution-mww]:

```{r, echo=FALSE, results='asis'}
show_prereg("ttest")
```

[^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes. The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.

We then translated this methods section into the following `planned_analysis()` function:

```{r planned_analysis}
```

This function illustrates two advantages of PAC.
First, a PAC can easily include data-dependent decisions creating different analysis branches under different conditions.
Second, it highlights how difficult it is to describe a statistical analysis precisely.
The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
Both are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
In contrast, the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
Still, these details can make a difference and provide unwanted researchers' degrees of freedom.

The function `planned_analysis()` together with the function `simulate_data()` (not shown here) can be used to justify the planned sample size.
To that end, `simulate_data()` is repeatedly called with increased sample size and the proportion of significant results (power) is recorded.
The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).

Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].

```{r, echo=FALSE}
res <-
  read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddddd")
choosen_power <- .8
choosen_d <- .2
minn <- filter(res, d == choosen_d, power > choosen_power) %>% 
  filter(n == min(n))
```

```{r, include=FALSE}
fig_caption <- str_c("Results of simulation for the power analysis. The cross indicates the sample size that archives ", choosen_power*100, "% assuming a Cohen's D of ", choosen_d, ".")
```

```{r power, echo=FALSE, fig.cap=fig_caption}
res %>%
  ggplot(aes(n, power, color = d, group = d)) +
  geom_line() +
  geom_point(
    data = minn,
    size = 3,
    color = "black",
    shape = 3
  ) +
  theme_minimal() +
  scale_color_viridis_c() +
  theme(legend.title = element_text()) +
  labs(y = "Power\n(Proportion of significant results)",
       x = "Sample size",
       color = "Cohen's D") +
  NULL
```

```{r, echo=FALSE, results='asis'}
show_prereg("power-code", wrap = FALSE)
```

```{r, echo=FALSE, results='asis'}
show_prereg("power")
```

We implemented a mechanism that only uses simulated data when the actual data is not yet available for the results section.
This mechanism also warns readers if these results are based on simulated data.
As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study.

```{r, echo=FALSE, results='asis'}
show_prereg("results", wrap = FALSE)
```

The last thing we need to do is to report our results with the combination of `planned_analysis()` and `report_analysis()`.

```{r, echo=FALSE, results='asis'}
show_prereg("report", wrap = FALSE)
```

This is an example of how the results could be reported (based on simulated data):

```{r, echo=FALSE, results='asis'}
set.seed(1234)
show(report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)), cat = FALSE), fence = "")
```

The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
<!-- I would show the rendered PDF (as long as it is shorter than one page) -->

# Discussion

<!--AB: TODO add a quick summary -->

# Limitations

PAC comes with a challenge.
In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
The reproducible workflow we present here is heavily automated and hence promotes reusability.
Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].

Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.

Please note that the statistical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and is not necessarily optimal for addressing your research question.
The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
<!-- nobody cares about limitations of the example analysis, I would delete all of this -->

We also suggest that version control with Git and GitHub fulfills the criteria of a preregistration as long as you state your research plan comprehensively.
This view is not shared by everyone [@szollosi_is_2020] because with some effort and malice intend the record of Git/GitHub can be altered.
This concern can be addressed by using zenodo.org to give a digital object identifier (DOI) to your preregistration (see [this guide](https://guides.github.com/activities/citable-code/)) because zenondo.org saves an unchangeable copy of your repository.

# Outlook

<!--AB: two lines of what the future of the package, of such standards, and prereigstration is. Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->

# References

```{r, include=FALSE}
knitr::write_bib(
  c(
    .packages(),
    "repro",
    "here",
    "rticles",
    "gert",
    "bookdown",
    "lavaan",
    "knitr",
    "targets",
    "renv",
    "tidyverse"
  ),
  here("packages.bib")
)
cat(c(readLines(here::here("packages.bib")), "\n", readLines(here::here("references.bib"))), sep = "\n", file = "temp.bib")
```
