---
title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
output: pdf_document
repro:
  packages:
    - usethis
    - repro
---

<!-- the HTML comments, like this one, are meta comments, mainly describing the intent --->
<!-- each sentence below a heading summarizes what I want to say there --->
<!-- "Hands-on:" means concrete practical application, they roughly proceed from easy/familiar to hard/unfamiliar-->

# Why you should care about reproducibility

<!-- define reproducibility -->

## Productivity

<!-- lure the reader in -->

Adherence to best practices promises to increase productivity for individual researchers and the community as a whole.

## Metascientific Arguments

<!-- present reproducibility as something that fits right into the researchers' objectives-->

Reproducibility is caused by openness, enables replication, and all three together warrant trust in science. 

# The lifecycle of an open research project

<!-- many small steps archive reproducibility -->

## Setup

<!-- keep it brief, let `repro` do the work -->

We assume that you have already installed R and RStudio (if not check the installation guide).
Additionally, you'll need the [`repro`-package](https://github.com/aaronpeikert/repro):

```r
if(!requireNamespace("remotes"))install.packages("remotes")
remotes::install_github("aaronpeikert/repro")
library("repro")
```

## Planing

<!-- researcher should begin early to enjoy most benefits -->

Organize files in one place.

<!-- Hands-on: Introducing RProjects-->

Keep notes.

<!-- Hands on: Introduce Markdown -->

Record changes.

<!-- Hands-on: Introduce Git -->

<!-- This is maybe the place to introduce the example-->
<!-- something roughly as complex as a bachelor thesis, like:

Hypothesis: Machivallianism is higher in male persons
Sample: Adults from western countries
Analytic strategy: Multigroup CFA + Partial Measurement Invariance
Data from: https://openpsychometrics.org/_rawdata/ either SD3 or MACH-IV
-->

```{r}
check_git()
```


### Simulation (optional)

<!-- impress with neatness → sample size planning, preregistration and analysis in one 😎-->
<!-- but explain that good science is often not as neat, strive for the ideal -->

Simulating data helps prevent unpleasant surprises and lets you double-check that the data you expect to gather and your analysis fit together.

<!--Hands-on: build simple functions tailored for your analysis-->

### Preregestration (optional)

A preregistration is a tool that may help to increase the credibility of empirical results.

<!--Hands on: Introduce RMarkdown-->
<!--Hands on: gitignore resulting .html/pdf-->
<!-- A preregistration needs to be public -->

## Prepare public release (optional)

<!--Hands on: Add a readme-->
```r
usethis::use_readme_rmd()
```

<!--Hands on: Add a license-->
```r
usethis::use_cc0_license()
```

<!--Hands on: Add a code of conduct-->
```r
usethis::use_code_of_conduct()
```

## Collaboration

### GitHub

GitHub/GitLab are great as a shared information basis: sharing and editing files, discussing and distributing tasks via issues and documenting the project via wikis even before data collection starts.

<!--Hands-on: Introduce Github-->

```{r}
check_github()
```

```r
usethis::use_github()
```

<!--Hands on: Introduce GitHub Issues-->
<!--Hands on: Introduce GitHub Projects-->

### Internal Dependencies

A well defined "entry point" that automatically reproduces everything eases collaboration.

<!--I am not sure this is the right place, should we introduce this earlier? -->

<!--Hands on: Introduce Make-->
<!--Hands on: Add targets for README.Rmd, preregistration.Rmd -->

``` r
automate_make()
```

### External Dependencies

A considerable number of external dependencies may hinder collaboration and long term reproducibility, but automated solutions can "bundle" and fix dependencies.

<!-- Hands on: Introduce Docker -->

``` r
automate_docker()
```

<-- after this everything is optional? -->

#### renv (alternative)

#### Singularity (optional)

## Analyzing Data

<!-- Hands on: add data via automate() -->
<!-- Hands on: automatically download data via manually editing the Makefile -->

### Data Integrety (optional)

<!-- Hands-on: check changed hash -->

### Anonymising Data (optional)

<!-- Should we add a section on synthetic data generation? -->
<!-- Hands on: Create synthetic data -->

## Releasing a Preprint (optional)

<!-- Hands on: add commit hash to results-->
<!-- Hands on: git tag + github release-->

## Peer Review (optional)

## Post Publication Review (optional)

## Archiving

<!-- Hands on: explain how to archive data-->
<!-- Hands on: explain how to archive docker images-->

