---
title: "Preregistration"
author: "Aaron Peikert"
date: "4/12/2021"
output: html_document
params:
  nsim: 2
  nmin: 100
  nmax: 300
  nstep: 100
  seed: 1234
repro:
  packages:
    - tidyverse
    - here
    - lavaan
    - furrr
    - future.batchtools
    - slider
    - patchwork
    - svglite
  scripts:
    - R/simulation.R
    - R/hash.R
---

```{r setup, include=FALSE}
library(tidyverse)
library(lavaan)
library(furrr)
library(slider)
library(patchwork)
source(here::here("R", "simulation.R"))
repro::automate_load_scripts()
# if you have access to a hpc envir specify this in R/hpc.R
# if not we use local multicore
# to speed up consider reduce nsim; increase nstep in ↑YAML↑

hpc_config <- here::here("R", "hpc.R")
if(fs::file_exists(hpc_config)){
  source(hpc_config)
} else {
  plan(list(transparent,
       tweak(multisession, workers = 4L)))
}
# debug:
#   plan(transparent)
```

```{r}
loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
cohend_jones_paulhus <- c(24, 29, 35)/100
```

```{r sim}
n_obs <- seq(params$nmin, params$nmax, params$nstep)
n_sim <- params$nsim
# n_sim * length(n_obs)
# [1] 100000
to_export <- ls_funs() %>% map(get)
to_export <-
  c(to_export, list(
    n_obs = seq(params$nmin, params$nmax, params$nstep),
    n_sim = params$nsim
  ))
res_raw %<-% simulation_study(n_obs, n_sim,
                furrr_options(
                  globals = to_export,
                  seed = seed,
                  packages = c("furrr", "lavaan", "tidyverse")
                ))
```

```{r write-out, error=TRUE, include=FALSE}
invisible(res_raw)
write_rds(res_raw, here::here("res_raw.rds"))
```

```{r hash}
```

```{r}
res <- res_raw %>% mutate(n_obs = as.numeric(n_obs))
res_interval <- res %>%
  select(n_obs, std_estimate) %>% 
  filter(!is.na(std_estimate)) %>% 
  group_by(n_obs) %>% 
  summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
            .groups = "drop", median = median(std_estimate)) %>% 
  unnest(c(interval)) %>% 
  mutate(width = upper - lower, perc = 1 - alpha) %>% 
  group_by(alpha) %>% 
  arrange(n_obs) %>% 
  mutate(across(c(width, upper, lower),
                ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE))) %>% 
  ungroup() %>% 
  arrange(perc)
```

```{r}
choosen_perc <- .8
choosen_width <- .2
true_diff <- .2
minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
  filter(n_obs == min(n_obs))
```

```{r}
plot_interval <- res_interval %>%
  ggplot(aes(
    n_obs,
    ymin = lower,
    ymax = upper,
    fill = perc,
    group = alpha
  )) +
  geom_ribbon() +
  geom_line(aes(y = median), color = "white") +
  geom_hline(yintercept = true_diff, linetype = "dotted") +
  scale_fill_viridis_c(
    begin = 1,
    end = 0,
    breaks  = 1 - seq(0.05, .2, 0.05),
    labels = . %>% scales::percent(accuracy = 1)
  ) +
  scale_y_continuous(n.breaks = 10) +
  scale_x_continuous(breaks = seq(0, params$nmax, params$nstep * 10)) +
  theme_minimal() +
  labs(x = "N per Group", y = "Standardized Effect Size") +
  theme(legend.title = element_blank()) +
  NULL
```

```{r}
plot_width <- res_interval %>%
  ggplot(aes(n_obs, width, color = perc, group = perc)) +
  geom_line() +
  geom_point(
    data = minn,
    size = 3,
    color = "black",
    shape = 3
  ) +
  scale_color_viridis_c(
    begin = 1,
    end = 0,
    breaks = 1 - seq(0.05, .2, 0.05),
    labels = . %>% scales::percent(accuracy = 1)
  ) +
  theme_minimal() +
  scale_y_log10(breaks = seq(0, 0.8, .1)) +
  scale_x_continuous(breaks = seq(0, params$nmax, params$nstep * 10)) +
  labs(x = "N per Group", y = "Width of quantile\n(Standardized Effect Size)") +
  theme(legend.title = element_blank()) +
  NULL
```

```{r}
simulation_plot <- plot_interval +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
  plot_width +
  plot_layout(nrow = 2, guides = 'collect') +
  plot_annotation(tag_levels = 'A')
simulation_plot
write_rds(simulation_plot, "simulation_plot.rds")
ggsave("simulation.svg", device = svglite::svglite, width = 7, height = 4)
```

Plot A depicts the range of effect sizes we observed in the simulation, where the colour encodes the density.
The light yellow interval shows where 80% of the estimates land.
The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes.

Plot B represents the width of the interval shown in Plot A on a logarithmic scale.
The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.

```{r}
nmax <- filter(res_interval, n_obs == max(n_obs))
bias <- (slice(nmax, 1)$median - true_diff)/true_diff
```

We can therefore conclude, that when we want to plan a study where we expect that `r scales::percent(choosen_perc)` of the estimates fall into a range of +/- `r choosen_width/2`, we need a minimum sample size of `r minn$n_obs` per group.
The simulation also shows that we overestimate the effect size by `r scales::percent(bias)` (true: `r true_diff`, median of observed: `round(slice(nmax, 1)$median, 2)`) with a sample size per group of `r slice(nmax, 1)$n_obs`.

# Session Info

```{r}
if(fs::file_exists(hpc_config)){
  list(local = sessioninfo::session_info(),
     login_node = value(future(sessioninfo::session_info())),
     worker = value(future(value(future(sessioninfo::session_info())))))
} else {
  sessioninfo::session_info()
}
```

