---
title: "A Preregistered Comparison of Male and Female Machiavellianism"
date: "7/10/2021"
author:
  - name: Aaron Peikert
    affil: 1, *
    orcid: 0000-0001-7813-818X
  - name: Caspar J. Van Lissa
    affil: 2, 3
    orcid: 0000-0002-0808-5024
  - name: Andreas M. Brandmaier
    affil: 1, 4
    orcid: 0000-0001-8765-6982
affiliation:
  - num: 1
    address: |
      Center for Lifespan Psychology---Max Planck Institute for Human Development
      Lentzeallee 94, 14195 Berlin, Germany
  - num: 2
    address: |
      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
  - num: 3
    address: |
      Open Science Community Utrecht, Utrecht, Netherlands
  - num: 4
    address: |
      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
      Berlin, Germany and London, UK
# firstnote to eighthnote
correspondence: |
  peikert@mpib-berlin.mpg.de
journal: psych
type: tutorial
status: submit
bibliography: preregistration.bib
abstract: |
  We test for group differences in machiavellism between self identified males and females in the MACH-IV in a well powered preregistered study.
output:
  bookdown::markdown_document2:
    base_format: rticles::mdpi_article
header-includes:
  -   \def\baselinestretch{0.98}
---

\setlength{\bibsep}{0cm}
\titlespacing{\section}{.1cm}{.1cm}{.1cm}

```{r setup, include=FALSE}
library(tidyverse)
library(report)
source(here::here("R", "simulation_funs.R"))
```

Theoretical Background
======================

There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self identified males and females. @muris2017 note that studies focusing on gender differences in machiavellianism are sparse. We therefore want to add to this body of literature.

Method
======

```{r, echo=FALSE}
res <-
  readr::read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddddd")
choosen_power <- .8
choosen_d <- .2
minn <- filter(res, d == choosen_d, power > choosen_power) %>% 
  filter(n == min(n))
```

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study [@simmons2012]. We use data available from [openpsychometrics.org](https://openpsychometrics.org/_rawdata/) from the online version of the MACH-IV [@christie1970]and included participants that have responded to at least one machiavellianism item and reported their gender as either "male" or "female". We conduct a t-test on the average of machiavellianism items, if the skew of this avarage is below 1.0, else we conduct a Mann--Whitney--Wilcoxon test between the gender groups. A simulation we conducted indicated that with a sample size of `r minn$n` for an alpha of .05 (two-sided) we archive at least `r choosen_power*100`% power assuming an effect size of `r choosen_d`.

Results
=======

```{r, echo=FALSE, results='asis'}
real_data <- here::here("data", "data.csv")
simulated <- !fs::file_exists(real_data)
if(simulated){
  cat("\\textcolor{red}{This document and the results presented herein are based on simulated data and should not be interpreted.}")
  set.seed(1234)
  mach <- simulate_data(900, 8, 0.3, 10)
} else {
  mach <- readr::read_csv(real_data)
}
```

```{r, echo=FALSE, results='asis'}
analysis <- planned_analysis(mach)
report <- function(analysis) {
  params <- report::report_parameters(analysis$test)
  table <- attributes(params)$table
  model <- report::report_model(analysis$test, table = table)
  if (analysis$use_rank)
    model <-
    str_replace(model,
                fixed("Welch Two Sample t-test"),
                "Mann--Whitney--Wilcoxon test")
  cat("The ", model, " suggests that the effect is ", params, sep = "")
}
```

```{r, echo=FALSE, results='asis'}
```

Discussion
==========

```{r, echo=FALSE, results='asis'}
if(simulated){
  cat("\\textcolor{red}{This document was created before we have gathered any data, we therefore have nothing to discuss.}")
}
```

References
==========
